---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Use unzip and read.csv to unpack and read data.

```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
Sum steps by date and make an histogram.
```{r histogram 1}
byDate <- aggregate(activity$steps, by=list(date = activity$date), FUN = sum, na.rm=TRUE, na.action=NULL)
names(byDate)[2] = "steps"
hist(byDate$steps, main="Histogram of Steps", xlab="Steps", breaks=10)
```

The mean of steps per day is `r format(mean(byDate$steps))`.

The median steps per day is `r format(median(byDate$steps))`.


## What is the average daily activity pattern?
Average steps per interval.
```{r time series}
byInterval <- aggregate(activity$steps, by=list(interval = activity$interval), FUN = mean, na.rm=TRUE, na.action=NULL)
names(byInterval)[2] = "avgSteps"
plot(byInterval$interval, byInterval$avgSteps, type="l", main="Daily activity", xlab= "Interval", ylab="Mean of Steps", col="blue")
```



## Imputing missing values
There are `r sum(!complete.cases(activity))` rows with NAs in the dataset.

Will use the average for each interval to fill in NAs.

```{r histogram 2}
merged <- merge(activity, byInterval, by = "interval")
merged$steps[is.na(merged$steps)] <- merged$avgSteps[is.na(merged$steps)]
byDate <- aggregate(merged$steps, by=list(date = activity$date), FUN = sum)
names(byDate)[2] = "steps"
hist(byDate$steps, main="Histogram of Steps (NAs replaced with interval mean)", xlab="Steps", breaks=10)
```

The mean of steps per day is `r format(mean(byDate$steps))`.

The median steps per day is `r format(median(byDate$steps))`.

"Do these values differ from the estimates from the first part of the assignment?"

**Yes.**
 
"What is the impact of imputing missing data on the estimates of the total daily number of steps?"

Intervals with fewer steps are better represented.
 
## Are there differences in activity patterns between weekdays and weekends?

Convert "date" column to date for use by function weekdays and classify weekdays.

```{r}
merged$date <- as.Date(merged$date)
weekend <- c('Saturday', 'Sunday')
merged$dayType <- factor((weekdays(merged$date) %in% weekend), levels=c(TRUE, FALSE), labels=c('weekend', 'weekday'))
```

Aggregated by interval and weekday.
```{r weekdays}
byIntervalWeekday <- aggregate(merged$steps, by=list(dayType=merged$dayType, interval = merged$interval), FUN = mean)
names(byIntervalWeekday)[3] = "avgSteps"
xyplot(avgSteps ~ interval | dayType, data = byIntervalWeekday, type = "l", main="Weekday vs weekend", ylab="Mean of Steps", layout=c(1,2))
```